---
title: "HW3-LogisticRegression"
author: "Richard Barad, Dave Drennan, Jarred Randall"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
library(aod)
library(rms)
library(gmodels)
library(nnet)
library(DAAG)
library(ROCR)
library(knitr)
library(kableExtra)
library(xtable)
library(dplyr)
library(tidyverse)
library(corrr)
library(crosstable)
library(caret)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
data <- read.csv("./Data/LogisticRegressionData.csv")

```

# Results

## Exploratory Data Analyis

```{r}

DRINKING_D.tab <- table(data$DRINKING_D)

kable(prop.table(DRINKING_D.tab), col.names = c("Drinking Driver", "Proportion"), 
      caption = "Proportion of Crashes Involving a Drinking Driver") %>%
  kable_styling(c("striped"), full_width = T, position = "left", font_size = 12)

```

## cross-tabulation between dependent variable and binary predictors

```{r crosstab, warning=FALSE, message=FALSE, cache=FALSE, results='hide'}

# We need to figure out the best way to make the table he is asking for

ct1_pct <- CrossTable(data$DRINKING_D, data$FATAL_OR_M, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct2_pct <- CrossTable(data$DRINKING_D, data$OVERTURNED, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct3_pct <- CrossTable(data$DRINKING_D, data$CELL_PHONE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct4_pct <- CrossTable(data$DRINKING_D, data$SPEEDING, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct5_pct <- CrossTable(data$DRINKING_D, data$AGGRESSIVE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct6_pct <- CrossTable(data$DRINKING_D, data$DRIVER1617, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]
ct7_pct <- CrossTable(data$DRINKING_D, data$DRIVER65PLUS, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$prop.row[3:4]

ct1_n <- CrossTable(data$DRINKING_D, data$FATAL_OR_M, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct2_n <- CrossTable(data$DRINKING_D, data$OVERTURNED, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct3_n <- CrossTable(data$DRINKING_D, data$CELL_PHONE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct4_n <- CrossTable(data$DRINKING_D, data$SPEEDING, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct5_n <- CrossTable(data$DRINKING_D, data$AGGRESSIVE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct6_n <- CrossTable(data$DRINKING_D, data$DRIVER1617, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
ct7_n <- CrossTable(data$DRINKING_D, data$DRIVER65PLUS, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE)$t[3:4]
```

```{r chi_squared, warning=FALSE, message=FALSE, cache=FALSE, results='hide'}
ct_chi1 <- CrossTable(data$FATAL_OR_M, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi2 <- CrossTable(data$OVERTURNED, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi3 <- CrossTable(data$CELL_PHONE, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi4 <- CrossTable(data$SPEEDING, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi5 <- CrossTable(data$AGGRESSIVE, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi6 <- CrossTable(data$DRIVER1617, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value
ct_chi7 <- CrossTable(data$DRIVER65PLUS, data$DRINKING_D, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE, chisq=TRUE)$chisq$p.value

chi_vector <- c(ct_chi1, ct_chi2, ct_chi3, ct_chi4, ct_chi5, ct_chi6, ct_chi7)

chi_table <- cbind(chi_vector,variable=c('FATAL_OR_M','OVERTURNED','CELL_PHONE','SPEEDING','AGGRESSIVE','DRIVER1617','DRIVER65PLUS')) %>%
  as_data_frame()
```

```{r}
df_cross_tab_pct <- cbind(DRINKING_D = c(0,1),ct1_pct, ct2_pct, ct3_pct, ct4_pct, ct5_pct, ct6_pct, ct7_pct) %>%
  as_data_frame() %>%
  rename(FATAL_OR_M = ct1_pct,
         OVERTURNED = ct2_pct,
         CELL_PHONE = ct3_pct,
         SPEEDING = ct4_pct,
         AGGRESSIVE = ct5_pct,
         DRIVER1617 = ct6_pct,
         DRIVER65PLUS = ct7_pct) %>%
  gather(variable, value, -DRINKING_D) %>%
  spread(key=DRINKING_D, value=value) %>%
  rename("No_Drinking_pct" = "0",
         "Drinking_pct" = "1")


cbind(DRINKING_D = c(0,1),ct1_n, ct2_n, ct3_n, ct4_n, ct5_n, ct6_n, ct7_n) %>%
  as_data_frame() %>%
  rename(FATAL_OR_M = ct1_n,
         OVERTURNED = ct2_n,
         CELL_PHONE = ct3_n,
         SPEEDING = ct4_n,
         AGGRESSIVE = ct5_n,
         DRIVER1617 = ct6_n,
         DRIVER65PLUS = ct7_n) %>%
  gather(variable, value, -DRINKING_D) %>%
  spread(key=DRINKING_D, value=value) %>%
  rename("No_Drinking_n" = "0",
         "Drinking_n" = "1") %>%
  left_join(.,df_cross_tab_pct,by='variable') %>%
  mutate(Total = No_Drinking_n + Drinking_n,
         Drinking_pct = round(Drinking_pct * 100,2),
         No_Drinking_pct = round(No_Drinking_pct * 100,2)) %>%
  select(variable, No_Drinking_n, No_Drinking_pct, Drinking_n, Drinking_pct,Total) %>%
  left_join(.,chi_table,by='variable') %>%
  kbl(col.names = c('Variable','N','%','N','%','Total','χ2 p-value')) %>%
  kable_classic_2() %>%
  add_header_above(header=c(" "=1,"No Alchoal Involved (DRINKING_D=0)"=2,"Alchoal Involved (DRINKING_D=1)"=2, " "=2))

```

## Calcuate Means for Continuous Variables

```{r}

data_group_mean <- data %>% dplyr::select(DRINKING_D,PCTBACHMOR,MEDHHINC) %>%
  group_by(DRINKING_D) %>% summarise_at(vars(PCTBACHMOR,MEDHHINC),mean) %>%
  gather(key='variable', value='mean', -DRINKING_D) %>%
  spread(key=DRINKING_D,value=mean) %>%
  rename("No_Drinking_mean" = "0",
         "Drinking_mean" = "1")

data_group_sd <- data %>% dplyr::select(DRINKING_D,PCTBACHMOR,MEDHHINC) %>%
  group_by(DRINKING_D) %>% summarise_at(vars(PCTBACHMOR,MEDHHINC),mean) %>%
  gather(key='variable', value='sd', -DRINKING_D) %>%
  spread(key=DRINKING_D,value=sd) %>%
  rename("No_Drinking_sd" = "0",
         "Drinking_sd" = "1")

```

```{r t_test}
t_PCTBACHMORE <- t.test(data$PCTBACHMOR~data$DRINKING_D)$p.value
t_MEDHHINC <- t.test(data$MEDHHINC~data$DRINKING_D)$p.value

t_vector = c(t_MEDHHINC,t_PCTBACHMORE)

cbind(data_group_mean,data_group_sd %>% dplyr::select(-variable),t_vector) %>%
  select(variable, Drinking_mean, Drinking_sd, No_Drinking_mean, No_Drinking_sd,t_vector) %>%
  kbl(col.names = c('Variable','Mean','SD','Mean','SD','t-test p-value')) %>%
  kable_classic_2() %>%
  add_header_above(header=c(" " = 1, "Drinking" = 2,"No Drinking"=2, " "=1))

```

## Multicollinearity checks

```{r pearson, echo=FALSE}

predictors <- data %>% dplyr::select(-CRN,-AREAKEY,-DRINKING_D)

predictors %>% 
  correlate() %>% 
  autoplot() +
  geom_text(aes(label = round(r,digits=2)),size = 3)
    
```

## Logistic Model

The table below shows the results of the logistic regression model which includes all nine predictors included. The statistically significant predictors are FATAL_OR_M, OVERTURNED, SPEEDING, AGGRESSIVE, DRIVER1617, and DRIVER65PLUS. We conclude these predictors are statistically significant because they have a p-value which is less than 0.05. The PCTBACHMOR and CELL_PHONE variables are not statistically significant.

```{r logit model}

logit = glm(formula = DRINKING_D ~ FATAL_OR_M + OVERTURNED + CELL_PHONE + SPEEDING + AGGRESSIVE + DRIVER1617 + DRIVER65PLUS + PCTBACHMOR + MEDHHINC, data = data, family = "binomial")

summary(logit)
```
We can also calculate the odds ratio for each Beta Coefficient. The Odds ratio is calculated by using the formula $e^{\beta_x}$. The calculations for the odds ratio are shown below and the odds ratios are also presented in the OR column of the table below.

* **FATAL_OR_M**: The odds ratio is equal to $e^{-0.814014} = 2.25694878$. This means that holding all other predictors constant, the odds of the driver being drunk increases by a factor of 2.25694878 when a fatality or major injury occurs during an accident.   

* **OVERTURNED**: The odds ratio is equal to $e^{-0.928914} = 2.53177687$. This means that holding all other predictors constant, the odds of the driver being drunk being drunk increases by a factor of 2.53177687 when a vehicle is overturned in an accident.

* **CELL_PHONE**: The odds ratio is equal to $e^{-0.02955008} = 1.02999102$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 1.02999102 when a driver was using a cell phone during the accident.

* **SPEEDING**: The odds ratio is equal to $e^{1.538976} = 4.65981462$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 4.65981462 when a driver was speeding during the accident. 

* **AGGRESSIVE**: The odds ratio is equal to $e^{-0.05969159} = 0.55050681$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 0.55050681 when a driver was aggressive during the accident. 

* **DRIVER1617**: The odds ratio is equal to $e^{-1.280296} = 0.27795502$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 0.27795502 when the crash involved at least one driver who is 16 or 17.  

* **DRIVER65PLUS**: The odds ratio is equal to $e^{-0.7746646} = 0.46085831$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 0.46085831 when the crash involved at least one driver who is 65 or older.

* **PCTBACHMORE**: The odds ratio is equal to $e^{-0.0003706336e-04} = 0.99962944$. This means that holding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 0.99962944 when the percent of 25+ year olds in the census block where the accident occurred goes up by 1%.

* **MEDHHINC**: The odds ratio is equal to $e^{00000.2804492} = 1.00000280$. This means that holiding all other predictors constant, the odds of the driver being drunk in an accident increases by a factor of 1.00000280 when the percent the median household income in the census block where the accident occured goes up by 1 USD. 

```{r}
logitoutput <- summary(logit)

or_ci <- exp(cbind(OR=coef(logit), confint(logit)))
or_ci

```


```{r warning=FALSE, message=FALSE, cache=FALSE}

fit <- logit$fitted       #Getting the y-hats (i.e., predicted values)
hist(fit)       #Histogram of fitted values
```
```{r}

tresholds = c(0.02,0.03,0.05,00.07,0.08,0.09,0.1,0.15,0.2,0.5)
sens = c()
spec = c()
mis_class = c()

for (x in tresholds){
  fit.binary = as.factor(ifelse(fit > x , 1, 0))
  results <- caret::confusionMatrix(fit.binary, as.factor(data$DRINKING_D), positive = "1")
  results_table <- results$table
  sens <- append(sens,results$byClass[1])
  spec <- append(spec,results$byClass[2])
  mis_class <- append(mis_class, 1 - results$overall[1])}

cbind(tresholds,sens,spec,mis_class) %>%
  as_data_frame() %>%
  kbl(col.names=c("Treshold","Sensitivity","Specificty","Miss Classification Rate")) %>%
  kable_classic_2()
```

## ROC Curve

```{r warning=FALSE, message=FALSE, cache=FALSE}
a <- cbind(data$DRINKING_D, fit)
head(a)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
colnames(a) <- c("labels","predictions")
head(a)
roc <- as.data.frame(a)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
pred <- prediction(roc$predictions, roc$labels)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
roc.perf = performance(pred, measure = "tpr", x.measure="fpr")
plot(roc.perf)
abline(a=0,b=1)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
opt.cut = function(perf, pred){
  cut.ind = mapply(FUN=function(x, y, p){
    d = (x - 0)^2 + (y-1)^2
    ind = which(d == min(d))
    c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
      cutoff = p[[ind]])
  }, perf@x.values, perf@y.values, pred@cutoffs)
}

print(opt.cut(roc.perf, pred))
```


```{r warning=FALSE, message=FALSE, cache=FALSE}
auc.perf = performance(pred, measure ="auc")
auc.perf@y.values
```

```{r logit model2}

logit2 = glm(formula = DRINKING_D ~ OVERTURNED + CELL_PHONE + SPEEDING + AGGRESSIVE + DRIVER1617 + DRIVER65PLUS, data = data, family = "binomial")

summary(logit2)

exp(cbind(OR = coef(logit2), confint(logit2)))

logitoutput2 <- summary(logit2)

logitcoeffs2 <- logitoutput2$coefficients
logitcoeffs2

or_ci2 <- exp(cbind(OR=coef(logit2), confint(logit2)))

finallogitoutput2 <- cbind(logitcoeffs2, or_ci2)
finallogitoutput2

```

```{r AIC}

AIC(logit, logit2)
```
